Relative Stable Maps in Genus One via Central Alignments

### 2019-01-01 GenusOneRelativeMaps.tex
Further improvements to Section 4

By pulling back the equation defining $H$ along $f$ and then taking its $\alpha_k$-th derivative at $x_k$ (which makes sense since all the lower-order derivatives vanish by assumption) we obtain a section of a jet bundle on $\VZ$ whose vanishing locus coincides (set-theoretically) with the locus where $x_k$ belongs to an internal component. As in \cite[Construction 2.1]{Ga} there is an exact sequence of jet bundles
\begin{equation*} 0\to x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to x_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to x_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0 \end{equation*}
and thus we obtain a section of the line bundle $x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)$ with vanishing locus equal to the locus where $x_k$ belongs to an internal component. Writing this as
\begin{equation*} \Dcal^k = \Dcal_{1,\alpha}^k(\PP^N|H,d) \subseteq \VZ_{1,\alpha}(\PP^N|H,d) = \VZ \end{equation*}
we obtain:
\begin{lem} $(\alpha_k\psi_k+\ev_k^*H)\cap[\VZ]=[\Dcal^k].\footnote{(Navid) Should we work out the vanishing orders of the section at this point?}$\end{lem}
\noindent The remainder of this section will be devoted to providing an explicit (recursive) description of $[\Dcal^k]$ in terms of tautological classes on spaces of maps with smaller numerical invariants. As we will see, this provides an algorithm for computing reduced relative invariants, as well as a quantum Lefschetz algorithm for reduced (absolute) invariants in genus one.

@@

We proceed as in the previous section, fixing an element of $\Mcal$ and considering the resulting subdivision of the tropical moduli space. As long as the minimal subcurve of genus one is not contracted, the picture is classical (i.e. the subdivision is trivial, and $\VZ=\Mcal$ locally). In this case there may only be one vertex of the tropical curve mapped into the interior of $\RR_{\geq 0}$; if there were more, the positions of their images in $\RR_{\geq 0}$ would be unconstrained and thus the corresponding maximal cone in the tropical moduli space would be at least $2$--dimensional, producing a locus in $\VZ$ of larger codimension.\footnote{(Luca) Say something about rigid tropical curves as in ACGS and KLR} We adopt the terminology of R. Vakil, distinguishing three cases as below:

@@

The description of the first three types is relatively easy:
\begin{enumerate}[label=(\alph*)]
 \item \[\mathcal Y_a=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_b=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_c^+= \VZc{1}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^{\sim} \times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
\end{enumerate}
where $\M{0}{\mu}{H}{d}^\sim$ denotes the space of genus zero punctured maps, while $\VZc{0}{\mu}{H}{d}$ denotes the desingularisation of the main component of the space of punctured maps discussed in Section ???. The numerical data run over all splitting $A$ of the tangency conditions (resp. $B$ of the degree) in $r+1$ parts, such that:
\begin{enumerate}
 \item $d_0+\sum_{i=1}^r m^{(i)}=\sum\alpha^{(0)}$;
 \item $d_i=m^{(i)}+\sum\alpha^{(i)}$;
 \item $d_0>0$ in case $c$.
\end{enumerate}

>>>>

By pulling back the equation defining $H$ along $f$ and then taking its $\alpha_k$-th derivative at $x_k$ (which makes sense since all the lower-order derivatives vanish by assumption) we obtain a section of a jet bundle on $\VZ$ whose vanishing locus coincides (set-theoretically) with the locus where $x_k$ belongs to an internal component. As in \cite[Construction 2.1]{Ga} there is an exact sequence of jet bundles
\begin{equation*} 0\to x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)\to x_k^*\mathcal J^{\alpha_k}(f^*\OO_{\PP^N}(H))\to x_k^*\mathcal J^{\alpha_k-1}(f^*\OO_{\PP^N}(H))\to 0 \end{equation*}
and thus we obtain a section of the line bundle $x_k^*\omega_{\mathcal{C}/\VZ}^{\otimes\alpha_k}\otimes\ev_k^*\OO_{\PP^N}(H)$ with vanishing locus equal to the locus where $x_k$ belongs to an internal component. Writing this as
\begin{equation*} \Dcal^k = \Dcal_{1,\alpha}^k(\PP^N|H,d) \subseteq \VZ_{1,\alpha}(\PP^N|H,d) = \VZ \end{equation*}
we obtain:
\begin{lem} $(\alpha_k\psi_k+\ev_k^*H)\cap[\VZ]=[\Dcal^k].\footnote{(Navid) Should we work out the vanishing orders of the section at this point?}$\end{lem}
\noindent The remainder of this section will be devoted to providing an explicit recursive description of $[\Dcal^k]$ in terms of tautological classes on spaces of maps with smaller numerical invariants. As we will see, this provides an algorithm for computing reduced relative invariants, as well as a quantum Lefschetz algorithm for reduced (absolute) invariants in genus one.

@@

We proceed as in the previous section, fixing an element of $\Mcal$ and considering the resulting subdivision of the tropical moduli space. In fact, it makes sense to choose a locally closed substack of $\Mcal$ where the combinatorial type of the tropical curve (i.e. the ghost sheaf) is constant. Then as long as the minimal subcurve of genus one is not contracted, the picture is classical (i.e. the subdivision is trivial, and $\VZ=\Mcal$ locally). In this case there can only be one vertex of the tropical curve mapped into the interior of $\RR_{\geq 0}$; if there were more, the positions of their images in $\RR_{\geq 0}$ would be independent and thus the corresponding maximal cone in the tropical moduli space would be at least $2$~dimensional, producing a locus in $\VZ$ of larger codimension.\footnote{(Luca) Say something about rigid tropical curves as in ACGS and KLR} We adopt the terminology of Vakil, distinguishing three cases as below:

@@

These divisors live inside an open neighbourhood of the preimage in $\VZ$ of our chosen locus in $\Mcal$, and are thus locally closed substacks of $\VZ$. To obtain the desired components of $\Dcal^k$, we must take the closure of these locally closed substacks. We will analyse this procedure on a case-by-case basis, and see once again that the tropical picture gives a complete and recursive description of the closure.

\subsubsection{Type A} By definition the locus $\Ycal_A$ contains those aligned maps which are in the closure of the locus of maps whose associated tropical curve has the same combinatorial type as Figure 1\footnote{(Navid) Bad LaTeX reference}. We will now argue that such an aligned map is always obtained by giving an alignment on the piece of the curve corresponding to the genus one vertex.

We start by considering possible limiting points of our original locus in $\Mcal$. These correspond to further degenerations of the (combinatorial type of the) tropical map.

\newpage
Now we may use the recursive description of the boundary in $\Mcal$ using punctured maps \cite{PuncturedMaps} to obtain a description of these loci:
\begin{enumerate}[label=(\Alph*)]
 \item \[\mathcal Y_a=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\VZc{1}{\alpha^{(1)}\cup\{m^{(1)}\}}{\PP^N|H}{d_1}\times\prod_{i=2}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_b=\M{0}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^\sim\times_{H^r}\left(\M{0}{\alpha^{(1)}\cup\{m^{(1)},m^{(2)}\}}{\PP^N|H}{d_1}\times\prod_{i=3}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\right)\]
 \item \[\mathcal Y_c^+= \VZc{1}{\{-m^{(1)},\ldots,-m^{(r)}\}\cup\alpha^{(0)}}{H}{d_0}^{\sim} \times_{H^r}\prod_{i=1}^r\M{0}{\alpha^{(i)}\cup\{m^{(i)}\}}{\PP^N|H}{d_i}\]
\end{enumerate}
where $\M{0}{\mu}{H}{d}^\sim$ denotes the space of genus zero punctured maps, while $\VZc{0}{\mu}{H}{d}$ denotes the desingularisation of the main component of the space of punctured maps discussed in Section ???. The numerical data run over all splitting $A$ of the tangency conditions (resp. $B$ of the degree) in $r+1$ parts, such that:
\begin{enumerate}
 \item $d_0+\sum_{i=1}^r m^{(i)}=\sum\alpha^{(0)}$;
 \item $d_i=m^{(i)}+\sum\alpha^{(i)}$;
 \item $d_0>0$ in case $c$.
\end{enumerate}
